Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19

### 2023-04-20 manuscript/manuscript.qmd
Audrey additions

AUDREY: Talk about the map of countries that derogated. The figure will go here-ish, I think

@@

During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21).

AUDREY: write expectations about what how derogating and non-derogating states differed in their likelihood of imposing emergency policies - canceling public events, gathering restrictions, closing public transit, movement and international travel

> **Hypothesis 1~a~**: If states blah, then blah.
>
> **Hypothesis 2~b~**: If states blah, then blah.
>

>>>>

Between January 2020 and June 2021, 23 states issued a total of 114 ICCPR derogations. [ INSERT MAP ABOUT HERE ] Of derogations that detailed specific obligation suspension lengths, the average length was 41 days. Descriptively, we see states across multiple regions derogated with notable concentration in South America, which has a robust history of engaging with human rights law. We also note that it is not necessarily the 'usual suspects' of strong democracies derogating.  Notably, Nordic states, which typically are very active in human rights law engagement did not issue any derogations during this period.

@@

During the pandemic, as @helfer2021 notes, states have had “sharply divergent responses” to international human rights law even when they are parties to the same treaty and confronting similar threats (21). As our descriptive data shows, not all states issued derogations. Derogating states were in the minority of states during the pandemic.

 We expect that states that derogate during COVID-19 with the intention to impose emergency policies will follow through. Our expectation comes from the specific, time-limited nature of derogations and the limited audience they speak to.  Critics of international law may expect states ratify treaties with the expectation of reputational and other benefits. However, derogations have no such public audience and are not signaling an increased commitment to human rights. There is no clear or expected connection with foreign aid, trade, or other theorized 'rewards' for larger legal actions. Instead, derogations are a very specific legal action that states voluntarily enter into after commitment has fully taken place.  Given this, we expect that derogating states took the legal action seriously and generally followed through with the derogations they submitted.

> **Hypothesis 1~a~**: If states issued COVID-19 derogations, then we expect that they will be more likely to impose relevant emergency polices such as canceling public events, restricting gatherings, closing public transit, and restricting domestic and international movement. 
>
> **Hypothesis 1~b~**: If states did not issue COVID-19 derogations, we expect that they will be less likely to impose related emergency policies.  
>
